---
title: "Tutorial for prospect"
author: "Jean-Baptiste Féret, Florian de Boissieu"
date: "`r Sys.Date()`"
output: 
  html_vignette:
    number_sections: true
vignette: >
  %\VignetteIndexEntry{Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval=FALSE
)
```


This tutorial briefly describes the different functionalities of the R package __prospect__ based on the leaf physical model.
PROSPECT aims at simulating leaf optical properties (directional-hemispherical reflectance and transmittance) in the optical domain from 400 nm to 2500 nm based on their biophysical properties, including a limited number of biochemical constituents and a unique structure parameter, N. 

The current version of the model implemented in __prospect__ is PROSPECT-PRO, which includes the following biochemical constituents:

* Chlorophyll a + b
* Carotenoids
* Anthocyanins
* Prown pigments
* Equivalent water thickness (EWT)
* Leaf mass per area (LMA) 
* proteins
* carbon-based consituents (CBC, constituents of dry matter other than proteins)

The specific absoprtion coefficients corresponding to these constituents are recorded in the variable `SpecPROSPECT` available when loading the package __prospect__. 

# Running __prospect__ in direct mode

## Input variables
The function `PROSPECT` runs PROSPECT for individual samples and expects the following input variables.

* `SpecPROSPECT`: dataframe including the refractive index and specific absorption ceofficients, defined for a given spectral range (max range: 400 nm - 2500 nm). simulation and inversion on different spctral domains can be peformed by adapting the information in `SpecPROSPECT`
* The biochemical and biophysical input variables of PROSPECT can then be introduced following two possible ways:

    * as individual values : these input variables include `N`, `CHL`, `CAR`, `ANT`, `BROWN`, `EWT`, `LMA`, `PROT`, `CBC`, `alpha`. If no value, their default value is set
    * `Input_PROSPECT`: a dataframe including the full list of PROSPECT input parameters defind earlier. 

## Output variables
`PROSPECT` returns a list containing directional-hemispherical reflectance and transmittance (`reflectance` and `transmittance`) corresponding to the input variables.


```{r prospect direct mode}
library(biodivMapR)
#############################################################
###         run PROSPECT using default parameters         ###
#############################################################
LRT   = PROSPECT(SpecPROSPECT)

#############################################################
###   run PROSPECT using user defind set of parameters    ###
#############################################################
LRT2   = PROSPECT(SpecPROSPECT,N = 1.4,CHL = 30,CAR = 10,EWT = 0.02,LMA = 0.01)
```


# Computing a Look-Up-Table with __prospect__

The function `PROSPECT_LUT` allows computation of a LUT dirctly based on a list of input parameters


```{r prospect LUT}
#############################################################
###         run PROSPECT in order to produce a LUT        ###
### if some parameters are not set, their default value   ###
###               will be added automatically             ###
#############################################################
CHL = 100*runif(1000)
CAR = 25*runif(1000)
ANT = 2*runif(1000)
EWT = 0.04*runif(1000)
LMA = 0.02*runif(1000)
N   = 1+2*runif(1000)
Input_PROSPECT  = data.frame('CHL'=CHL,'CAR'=CAR,'ANT'=ANT,'EWT'=EWT,'LMA'=LMA,'N'=N)
LUT             = PROSPECT_LUT(SpecPROSPECT,Input_PROSPECT)
```

