---
title: "1. Running PROSPECT in forward mode"
author: "Jean-Baptiste Féret, Florian de Boissieu"
date: "`r Sys.Date()`"
output:
  html_vignette:
    number_sections: true
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 2
vignette: >
  \usepackage[utf8]{inputenc}
  %\VignetteIndexEntry{1. Forward mode}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval=FALSE
)
```

<style>
  .col2 {
    columns: 2 180px;         /* number of columns and width in pixels*/
    -webkit-columns: 2 180px; /* chrome, safari */
    -moz-columns: 2 180px;    /* firefox */
  }
</style>


# Running `prospect` in forward mode

### Input variables
The function `prospect` runs the model PROSPECT to simulate leaf 
directional-hemispherical reflectance and transmittance from a set of chemical 
constituents and a structure parameter. The leaf optical properties are 
simulated over a spectral domain defined between 400 nm and 2500 nm, with a 
maximum spectral sampling of 1 nm.  

`prospect` uses the following input variables: 

* `spec_prospect`: a data frame including the refractive index and specific 
absorption coefficients (which are optical constants for a given version of PROSPECT).
** `spec_prospect` also defines the spectral domain used for simulation. 
** The default value of `spec_prospect` is defined by `prospect::spec_prospect_full_range`, 
which includes optical constants over the spectral domain ranging from 400 nm 
to 2500 nm. 
** Simulation and inversion on different spectral domains (reduced spectral 
range and/or adjusted spectral sampling) can be performed by adapting the 
information in `spec_prospect`.


* The biochemical and biophysical input variables of PROSPECT are :

<div class="col2">

  *  `n_struct` (default = 1.5)
  *  `chl` (default = 40.0  $\mu g.cm^2$)
  *  `car` (default = 8.0  $\mu g.cm^2$)
  *  `ant` (default = 0.0  $\mu g.cm^2$)
  *  `brown` (default = 0.0 arbitrary units)
  *  `ewt` (default = 0.01  $g.cm^2$)
  *  `lma` (default = 0.008  $g.cm^2$)
  *  `prot` (default = 0.0  $g.cm^2$)
  *  `cbc` (default = 0.0  $g.cm^2$)
  *  `alpha` (default = 40.0 degrees)

</div>


### Output variables
The `prospect` function returns a data frame containing leaf 
directional-hemispherical reflectance and transmittance (`reflectance` and 
`transmittance`) and the corresponding wavelengths (`wvl`) as defined in 
`spec_prospect`.

### run PROSPECT using default parameters over the full range from 400 nm to 2500 nm

When run without input parameters, `prospect` is run with default values: 
* `spec_prospect` is defined by `prospect::spec_prospect_full_range`
* the default values of the structural and chemical parameters are set. 

```{r prospect direct mode default}
library(prospect)
lrt_default <- prospect()
# the following command is equivalent when simulating over the full spectral domain 
lrt_default <- prospect(spec_prospect = prospect::spec_prospect_full_range)

```

###   run PROSPECT with user defined parameters over the 400-2500 nm range
Part or all of the input parameters can be provided as input parameters. 
Default values are set for undeclared parameters. 
This example simulates leaf optics over the full visible to shortwave infrared 
domain defined in PROSPECT. 

```{r prospect direct mode VSWIR}
lrt_vswir <- prospect(n_struct = 1.4, chl = 30, car = 6, ewt = 0.02, lma = 0.01)
```

###   run PROSPECT with user defined spectral domain
The spectral range in `spec_prospect$lambda` defines the simulated spectral domain.
It can be adjusted with the function `fit_spectral_data`.

`fit_spectral_data` can also be used to match the spectral range and sampling of 
input data with those of the optical constants. 

```{r prospect direct mode VNIR}
# define the spectral range for simulations in the VNIR from 400 to 1000 nm
wl_range <- seq(400,1000)
# adjust spectral properties used in PROSPECT to VNIR domain
adjust_vnir <- fit_spectral_data(lambda = wl_range)
lrt_vnir <- prospect(spec_prospect = adjust_vnir$spec_prospect,
                     n_struct = 1.4, chl = 30, car = 6, ewt = 0.02, lma = 0.01)
```

# Comparison between PROSPECT-PRO and PROSPECT-D

PROSPECT-D is the most commonly used version of PROSPECT, combining all dry 
matter constituents in a unique chemical constituent.
The absorption of this chemical constituent is accounted for through the variable `lma`. 

PROSPECT-PRO is the latest official version released. 
The only difference between PROSPECT-PRO and PROSPECT-D is that LMA is divided 
into proteins and CBC. Therefore, the default values in `prospect` correspond to 
calling PROSPECT-D, as `prot` and `cbc` are set to 0. 
When calling `prospect`, either `lma` or `prot` and `cbc` should be set to 0, 
or undefined. 
A message will be displayed if it is not the case:

```{r message mix PROSPECT-D and PROSPECT-PRO}
lrt_confusion_version <- prospect(chl = 45, car = 10, ant = 0.2, ewt = 0.012, 
                                  lma = 0.01, prot = 0.001, n_struct = 1.3)
'prot and/or cbc are not set to 0
lma is not set to 0 neither, which is physically incorrect
(lma = prot + cbc)
We assume that PROSPECT-PRO was called and set lma to 0
Please correct input parameters lma, prot and/or cbc if needed'
```

#### PROSPECT-D

Here is an example to run PROSPECT-D:

```{r run PROSPECT-D}
lrt_d <- prospect(chl = 45, car = 10, ant = 0.2, 
                  ewt = 0.012, lma = 0.01, n_struct =1.3)
```

and another one with PROSPECT-PRO, which should lead to very similar leaf optics: 

```{r run PROSPECT-PRO}
lrt_pro <- prospect(chl = 45, car = 10, ant = 0.2, ewt = 0.012, 
                    prot = 0.001, cbc = 0.009, n_struct = 1.3)
```

The resulting leaf optical properties are, indeed, very similar:

<p>&nbsp;</p>
```{r, eval=TRUE, echo = FALSE, out.width="70%"}
url <- "https://gitlab.com/jbferet/myshareddata/-/raw/master/prospect_vignette_illustrations/compare_PROSPECT.png"
knitr::include_graphics(url)
```
<p>&nbsp;</p>
<center>
  Fig. 1. Comparison between PROSPECT-D and PROSPECT-PRO
</center> 
<p>&nbsp;</p>


while using proteins only instead of proteins + CBC to simulate LMA leads to 
different results: 

```{r run PROSPECT-PRO2}
lrt_pro2 <- prospect(chl = 45, car = 10, ant = 0.2, ewt = 0.012, 
                     prot = 0.010, cbc = 0.0, n_struct = 1.3)
```

<p>&nbsp;</p>
```{r, eval=TRUE, echo = FALSE, out.width="70%"}
url <- "https://gitlab.com/jbferet/myshareddata/-/raw/master/prospect_vignette_illustrations/compare_PROSPECT2.png"
knitr::include_graphics(url)
```
<p>&nbsp;</p>
<center>
  Fig. 2. Comparison between PROSPECT-D and PROSPECT-PRO, Proteins only, no CBC
</center> 
<p>&nbsp;</p>


On the other hand, using CBC only instead of proteins + CBC to simulate LMA 
leads to very similar results compared to simulation with PROSPECT-D. 

This is explained by the low proportion of proteins compared to CBC in the total 
contribution to LMA, and the very similar specific absorption coefficient between 
LMA and CBC. This also highlights the challenges for the proper estimation of 
proteins from LOP.

```{r run PROSPECT-PRO3}
lrt_pro3 <- prospect(chl = 45, car = 10, ant = 0.2, ewt = 0.012, 
                     prot = 0.000, cbc = 0.010, n_struct = 1.3)
```


<p>&nbsp;</p>
```{r, eval=TRUE, echo = FALSE, out.width="70%"}
url <- "https://gitlab.com/jbferet/myshareddata/-/raw/master/prospect_vignette_illustrations/compare_PROSPECT3.png"
knitr::include_graphics(url)
```
<p>&nbsp;</p>
<center>
  Fig. 3. Comparison between PROSPECT-D and PROSPECT-PRO, CBC only, no Proteins
</center> 
<p>&nbsp;</p>


### Run previous versions from PROSPECT-PRO

If you want to run previous versions of PROSPECT:

*  __We do not recommend using PROSPECT-4 or PROSPECT-5__, as the calibration of 
the refractive index along with specific absorption coefficients for chlorophylls 
and carotenoids resulted in artifacts in the VIS domain. 

*  If you want to account for the absorption of chlorophylls and carotenoids, 
but do not want to include anthocyanins (as it is the case for PROSPECT-5): 
calling `prospect` with no input `ant` will set its default value to 0. 

*  If you do not want to differentiate chlorophylls and carotenoids  (as it is 
the case for PROSPECT-4 and earlier versions): we recommend that you use 
PROSPECT and set `car` as a constant fraction of `chl`.

# Computing a Look-Up-Table with `prospect`

Look-Up-Tables (LUT) are widely used in order to infer leaf characteristics from 
PROSPECT, based on minimization techniques. The function `prospect_lut` allows 
computation of a LUT directly based on a data frame of input parameters.

The following example produces LUTs in the VSWIR and VNIR domains with the 
function `prospect_lut` of `prospect`. 
Undefined parameters will be automatically set to their default value. 

```{r prospect LUT}
input_prospect <- data.frame('chl' = 100*runif(1000), 
                             'car' = 25*runif(1000), 
                             'ant' = 2*runif(1000), 
                             'ewt' = 0.04*runif(1000), 
                             'lma' = 0.02*runif(1000), 
                             'n_struct' = 1+2*runif(1000))
# produce a LUT defined over the VSWIR domain covered by PROSPECT
lut <- prospect_lut(input_prospect = input_prospect)

# produce a LUT defined over the VNIR domain defined previously
lut_vnir <- prospect_lut(spec_prospect = adjust_vnir$spec_prospect,
                         input_prospect = input_prospect)

```
