---
title: "2. Running PROSPECT in inverse mode: iterative optimization"
author: "Jean-Baptiste Féret, Florian de Boissieu"
date: "`r Sys.Date()`"
output:
  html_vignette:
    number_sections: true
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: false
    toc_depth: 2
vignette: >
  \usepackage[utf8]{inputenc}
  %\VignetteIndexEntry{2. Inverse mode: iterative optimization}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval=FALSE
)
```

# Running `prospect` in inverse mode: generalities

`PROSPECT` can be inverted using iterative optimization, by calling the function `Invert_PROSPECT`. 
This iterative optimization is based on the function [`fmincon` included in the package `pracma`](https://rdrr.io/cran/pracma/man/fmincon.html).

By default, the merit function used for the inversion minimizes the RMSE between the simulated and the measured leaf optical properties.
However, users can define their own merit function with associated criterion to minimize by defining their own merit function and adding it as input variable, such as `MeritFunction = MyOwnMeritFunction`.

The results of inversion will also depend on the prospect version, i.e. on the parameters included in the prospect model. 
The following table shows the correspondence between prospect versions and prospect parameters. 
Each parameter can be considered as parameter to estimate or a fix value in the model, depending on the prior knowledge. 

| Version  | 5                                       | 5B                                    | D                                      | DB                                     | PRO                                    | PROB
| :------: |:--------------------------------------:|:--------------------------------------:|:--------------------------------------:|:--------------------------------------:|:--------------------------------------:|:---------------------------------------:|
| CHL      |`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|
| CAR      |`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|
| ANT      |                                        |                                        |`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|
| BROWN    |                                        |`r emojifont::emoji('white_check_mark')`|                                        |`r emojifont::emoji('white_check_mark')`|                                        |`r emojifont::emoji('white_check_mark')`|
| EWT      |`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|
| LMA      |`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|                                        |
| PROT     |                                        |                                        |                                        |                                        |`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|
| CBC      |                                        |                                        |                                        |                                        |`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|
| N        |`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|`r emojifont::emoji('white_check_mark')`|

## Input variables
The function `Invert_PROSPECT` requires either reflectance, or transmittance, or both.
User can define which input variables of `PROSPECT` should be estimated during inversion, and which ones should be set to a given value (0 or user's choice).
The list of input variables for inversion is :

* `SpecPROSPECT`: data frame including the spectral bands, refractive index and specific absorption coefficients. 
`SpecPROSPECT = NULL` uses the spectral domain from 400 nm to 2500 nm. 
Simulation and inversion on different spectral domains can be performed by adjusting `SpecPROSPECT`
* `Refl`: numeric: individual leaf reflectance corresponding to the spectral 
domain defined in `SpecPROSPECT`. Set to `NULL` if inversion on transmittance only
* `Tran`: numeric: individual leaf transmittance corresponding to the spectral 
domain defined in `SpecPROSPECT`. Set to `NULL` if inversion on reflectance only
* `Parms2Estimate` list. Parameters to estimate. Set to 'ALL' by default. 
* `InitValues` data frame including initial values of PROSPECT input parameters. 
During optimization, they are used either as initialization values for parameters 
to estimate, or as fix values for other parameters. 
Parameters are not taken into account when not compatible with PROSPECT_version.
* `PROSPECT_version` character. Corresponds to the PROSPECT version. 
Should be one of the following versions: 'D', 'PRO'. 
* `MeritFunction` character. name of the function to be used as merit function 
with given criterion to minimize (default = Merit_PROSPECT_RMSE)
* `xlub` data frame. Boundaries of the parameters to estimate. 
The data frame must have columns corresponding to \code{Parms2Estimate} first 
line being the lower boundaries and second line the upper boundaries.
* `Est_Brown_Pigments` boolean. Should `BROWN` be estimated or not? Brown pigments are found during later stages of senescent leaves, but they are usually not found in juvenile, mature or early senescent leaves. 
* `Est_alpha` boolean. Should `alpha` be estimated or not? Keep in mind that most published results use `alpha` with its default value.

## Output variables
`Invert_PROSPECT` returns a list of estimated PROSPECT parameters defined in `Parms2Estimate`.


# Running PROSPECT-D inversion 

## Over the full spectral domain
All parameters are estimated, except `alpha` and `BROWN` which are set to their default value.
Brown pigments are characteristic of late senescent stages. 
They are absent for all other development stages, hence the assessment of brown pigments is not systematic here. 

```{r prospect inverse mode 1}
# simulate leaf optical properties
LRT_D <- PROSPECT(CHL = 45, CAR = 10, ANT = 0.2, 
                  EWT = 0.012, LMA = 0.01, N = 1.3)
# define set of parameters to be estimated
Parms2Estimate  <- 'ALL'
# invert PROSPECT with simulated leaf optical properties
OutPROSPECT <- Invert_PROSPECT(Refl = LRT_D$Reflectance, 
                               Tran = LRT_D$Transmittance,
                               Parms2Estimate = Parms2Estimate, 
                               PROSPECT_version = 'D')


# simulate leaf optical properties with brown pigments
LRT_D_brown <- PROSPECT(CHL = 45, CAR = 10, ANT = 0.2, BROWN = 0.1, 
                        EWT = 0.012, LMA = 0.01, N = 1.3)
# define set of parameters to be estimated
Parms2Estimate  <- 'ALL'
# invert PROSPECT with simulated leaf optical properties and specify that brown 
# pigment should be assessed in addition to all other constituents
OutPROSPECT <- Invert_PROSPECT(Refl = LRT_D_brown$Reflectance, 
                               Tran = LRT_D_brown$Transmittance,
                               Parms2Estimate = Parms2Estimate, 
                               PROSPECT_version = 'D', 
                               Est_Brown_Pigments = TRUE)
```


## over the VNIR domain
All parameters are estimated, except `alpha` and `BROWN` which are set to their default value.
`FitSpectralData` directly adjusts reflectance, transmittance and optical constants to user defined spectral bands. 
These spectral bands should be integer values between 400 and 2500, and can be continuous or not. 

```{r prospect inverse mode 2}
# define spectral subdomain from 400 nm to 800 nm
SpectralSubDomain <- seq(400,800)
# adjust spectral domain for reflectance, transmittance and SpecPROSPECT
SubData <- FitSpectralData(lambda = LRT_D$wvl, 
                           Refl = LRT_D$Reflectance, 
                           Tran = LRT_D$Transmittance,
                           UserDomain = SpectralSubDomain)

# Note that FitSpectralData can also run with UserDomain defining 
# upper and lower bounds when setting input argument 'UL_Bounds = TRUE'
ULBounds <- c(400,800)
SubData <- FitSpectralData(lambda = LRT_D$wvl, 
                           Refl = LRT_D$Reflectance, 
                           Tran = LRT_D$Transmittance,
                           UserDomain = ULBounds, 
                           UL_Bounds = TRUE)

# invert PROSPECT with simulated leaf optical properties
OutPROSPECT <- Invert_PROSPECT(SpecPROSPECT = SubData$SpecPROSPECT, 
                               Refl = SubData$Refl, 
                               Tran = SubData$Tran,
                               Parms2Estimate = Parms2Estimate, 
                               PROSPECT_version = 'D')
```

## over the VNIR domain with LMA and EWT value set
Only pigments and `N` are estimated. 
The same optical properties as previous example are used.

```{r prospect inverse mode 3}
# define set of parameters to be estimated
Parms2Estimate <- c('CHL', 'CAR', 'ANT', 'N')
# define set of parameters to be estimated
InitValues <- c('EWT' = 0.01, 'LMA' = 0.01)
# invert PROSPECT with simulated leaf optical properties
OutPROSPECT <- Invert_PROSPECT(SpecPROSPECT = SubData$SpecPROSPECT, 
                               Refl = SubData$Refl, 
                               Tran = SubData$Tran,
                               Parms2Estimate = Parms2Estimate, 
                               PROSPECT_version = 'D', 
                               InitValues = InitValues)
```

##   Run PROSPECT-D inversion over the SWIR domain between 1700 nm and 2400 nm
`EWT`, `LMA` and `N` are estimated. 
The same optical properties as previous example are used.

```{r prospect inverse mode 4}
# define set of parameters to be estimated
Parms2Estimate <- c('EWT', 'LMA', 'N')
# define spectral subdomain 
SpectralSubDomain <- seq(1700,2400)
# adjust spectral domain
SubData <- FitSpectralData(lambda = LRT_D$wvl, 
                           Refl = LRT_D$Reflectance, 
                           Tran = LRT_D$Transmittance,
                           UserDomain = SpectralSubDomain)

# invert PROSPECT with simulated leaf optical properties
OutPROSPECT <- Invert_PROSPECT(SpecPROSPECT = SubData$SpecPROSPECT, 
                               Refl = SubData$Refl, 
                               Tran = SubData$Tran,
                               Parms2Estimate = Parms2Estimate, 
                               PROSPECT_version = 'D')
```

#   Invert PROSPECT-D using optimal spectral domains
The function `Invert_PROSPECT_OPT` automatically sets the optimal spectral domains during inversion for all constituents to be estimated.

Optimal spectral domains and configuration are defined in [Féret et al. (2019)](https://doi.org/10.1016/j.rse.2018.11.002), [Féret et al. (2021)](https://doi.org/10.1016/j.rse.2020.112173), and [Spafford et al. (2021)](https://doi.org/10.1016/j.rse.2020.112176). 

`N` does not need to be part of `Parms2Estimate`, as it is automatically estimated when needed. 


```{r prospect inverse mode 5}
# define set of parameters to be estimated
Parms2Estimate  <- c('CHL','CAR','ANT','EWT','LMA')
# call Invert_PROSPECT_OPT to get optimal estimation of leaf parameters 
# using optimal spectral domains
ParmEst <- Invert_PROSPECT_OPT(lambda = LRT_D$wvl, 
                               Refl = LRT_D$Reflectance,
                               Tran = LRT_D$Transmittance,
                               PROSPECT_version = 'D',
                               Parms2Estimate = Parms2Estimate)
```

#   run PROSPECT-PRO inversion
Such definition of optimal spectral domains can also be set manually. 
For example, here is how to estimate protein content from leaf optical properties using the optimal spectral domain defined in [Féret et al. (2021)](https://doi.org/10.1016/j.rse.2020.112173).

Please note that `N` needs to be added to `Parms2Estimate`, if user want it to be estimated during the inversion, otherwise it will be set to its default value.


```{r prospect inverse mode 6}
# simulate leaf optical properties
LRT_PRO <- PROSPECT(CHL = 45, CAR = 10, ANT = 0.2, EWT = 0.012,
                    PROT = 0.002, CBC = 0.015, N = 1.3)
# define spectral subdomain 
SpectralSubDomain <- c(2125,2175)
# adjust spectral domain
SubData <- FitSpectralData(lambda = LRT_PRO$wvl, 
                           Refl = LRT_PRO$Reflectance, 
                           Tran = LRT_PRO$Transmittance,
                           UserDomain = SpectralSubDomain, 
                           UL_Bounds = TRUE)

Parms2Estimate <- c('EWT', 'PROT', 'CBC', 'N')
# invert PROSPECT with simulated leaf optical properties
OutPROSPECT <- Invert_PROSPECT(SpecPROSPECT = SubData$SpecPROSPECT,
                               Refl = SubData$Refl, 
                               Tran = SubData$Tran, 
                               Parms2Estimate = Parms2Estimate, 
                               PROSPECT_version = 'PRO')
```


# Run PROSPECT inversion with user-defined merit function

A new merit function can be defined, as long as it uses the same input and output
variables as the function [Merit_PROSPECT_RMSE](https://jbferet.gitlab.io/prospect/reference/Merit_PROSPECT_RMSE.html).

Here is an example of a function using __nMRSE__ instead of __RMSE__ for the 
merit function.
This means that for each spectral band, the square difference between measured and 
simulated leaf optics is normalized by the value of the measured leaf optics. 

First, the merit function is 

```{r prospect merit function}
Merit_PROSPECT_nRMSE <- function(x, SpecPROSPECT,
                                 Refl, Tran,
                                 Input_PROSPECT,
                                 Parms2Estimate) {
  x[x < 0] <- 0
  Input_PROSPECT[Parms2Estimate] <- x
  RT <- do.call('PROSPECT', c(list(SpecPROSPECT = SpecPROSPECT), Input_PROSPECT))
  fcr <- fct <- 0
  if (!is.null(Refl)) fcr <- sqrt(sum(((Refl - RT$Reflectance)**2)/Refl) / length(RT$Reflectance))
  if (!is.null(Tran)) fct <- sqrt(sum(((Tran - RT$Transmittance)**2)/Tran) / length(RT$Transmittance))
  fc <- fcr + fct
  return(fc)
}
```

```{r prospect inverse mode merit function}
# simulate leaf optical properties
LRT <- PROSPECT(CHL = 45, CAR = 10, ANT = 0.2, 
                EWT = 0.02, LMA = 0.015, N = 1.8)
Parms2Estimate <- c('CHL', 'CAR', 'ANT', 'EWT', 'LMA', 'N')
# invert PROSPECT with simulated leaf optical properties
OutPROSPECT <- Invert_PROSPECT(Refl = LRT$Refl, 
                               Tran = LRT$Tran, 
                               Parms2Estimate = Parms2Estimate, 
                               PROSPECT_version = 'D', 
                               MeritFunction = 'Merit_PROSPECT_nRMSE')
```

